---
title: "Lecture 4"
author: "Kamal"
date: "2023-09-29"
output: pdf_document
---

# My file

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

## Demo

### 0. Load the `tidyverse` Package

This section loads the packages we need in this lecture.

```{r}
library(tidyverse)
```

### 1. Import and Export the V-Dem Data

This section loads the VDEM dataset and describe its basic information

```{r}
d <- read_csv("/Users/kamalpathak/Desktop/Personal R /POLI3148_DaSPPA_Portfolio/_DataPublic_/vdem/1984_2022/vdem_1984_2022_external.csv")

```

### 2. Select economic development indicators

First, we look at the identifiers of the data.

```{r}
names(d)
```

```{r}
d |> # |> is officially called "pipe" it means passing down whatever in the previous step to the next step
  select(country_name, country_id, year) |>
  distinct() # distinct get rid of repetitions
```

```{r}
# Which countries are in this dataset
# Unique list of country names
d |> select(country_name) |> distinct()
```

```{r}
d |> select(year) |> distinct()
```

Select both the country identifiers, GDP, and GDP per capita.

```{r}
# cmd + shift + return runs everything within the chunk.
# <- is "assign" - create a new dataframe.
d_gdp <- d |> 
  select(country_name, country_id, year, e_gdp, e_gdppc)

d_gdp
```

### 3. Rename Columns to Make Names Informative

```{r}
# Rename columns
# rename("new name" = "old name")
d_gdp <- d_gdp |>
  rename("GDP" = "e_gdp", "GDP_per_capita" = "e_gdppc",
         "Country" = "country_name", "ID" = "country_id",
         "Year" = "year")

d_gdp # only for printing in Rmarkdown

# # You can create new dataframe in step
# d_gdp_2 <- d_gdp |>
#   rename("GDP" = "e_gdp", "GDP_per_capita" = "e_gdppc",
#          "Country" = "country_name", "ID" = "country_id",
#          "Year" = "year")
```

### 4. `slice` Rows

```{r}
# Want countries-years with highest GDP
d_gdp |> 
  slice_max(order_by = GDP, n = 10)
```

```{r}
# Get countries-years with the lowest GDP
d_gdp |> slice_min(order_by = GDP, n = 10)
```

```{r}
set.seed(52)
d_gdp |> slice_sample(n = 10) # Sample 10 observations
```

```{r}
set.seed(52)
d_gdp |> slice_sample(prop = 0.1)
```

### 5. Subset data by row

```{r}
# Want: 2000-2005 data
d_gdp |> filter(Year >= 2000 & Year <= 2008)
```

```{r}
d_gdp |> filter(Country == "China")
```

```{r}
# Want: 2000 - 2005 from China
d_gdp |> 
  filter(Year >= 2000 & Year <= 2005) |> 
  filter(Country == "China")
```

### 6. Arrange

```{r}
# Want: sort the row by GDP per capita
d_gdp |> arrange(GDP_per_capita) # smallest to largest
```

```{r}
d_gdp |> arrange(-GDP_per_capita) # largest to smallest. minus sign
```

### 7, Perform (4) (5) (6) group by group: group_by, ungroup

```{r}
# We want the highest gdp for each country
d_gdp |>
  group_by(Country) |>
  slice_max(GDP, n=1)
```

```{r}
# How many entries for each country
d_gdp |>
  group_by(Country) |>
  count()


```

```{r}
# Worst GDP for each country, get the year
d_gdp |> 
  group_by(Country) |>
  slice_min(order_by = GDP, n=1)
  
```

```{r}
# Best GDP
d_gdp |> 
  group_by(Country) |>
  slice_max(order_by = GDP, n=1)
  
```

### 8, Create new columns using mutate, group_by

```{r}
d_gdp |>
  mutate(New=1)

d_gdp |>
  mutate(New=GDP)

d_gdp |>
  mutate(New=log(GDP))

# Want new column to be GDP relative to average GDP in the world ( 1984-2022)
d_gdp |>
  mutate(GDP_over_average = GDP / mean(GDP, na.rm=TRUE))

# Want new column to be GDP relative to average GDP of the country in the world ( 1984-2022)

d_gdp |>
  group_by(Country)|>
  mutate(GDP_over_average = GDP / mean(GDP, na.rm=TRUE))


```

```{r}
# Country-year development level with reference to that of 1984.
d_gdp |>
  group_by(Country) |>
  arrange(Year) |>
  mutate(GDP_over_1984 = GDP / first(GDP)) |>
  ungroup() |>
  arrange(Country, Year)
```

```{r}
# Year on Year Economic Growth

d_gdp |> 
  group_by(Country) |>
  arrange(Year) |>
  mutate(GDP_yoy_change = GDP - lag(GDP, n=1)) |>
  ungroup() |>
  arrange(Country, Year)
```

### 9, Summarize the data : group_by, summarise, ungroup

```{r}
# Want: Average GDP level of the world

d_gdp |>
  summarise(
    gdp_average = mean(GDP, na.rm=TRUE),
    gdp_per_capita_average = mean(GDP_per_capita, na.rm=TRUE)
    )


```

```{r}
#Data Availability and  Integrity

d_gdp |> 

  #Craete a column that indicates whether the value is missing
  mutate(GDP_missing = as.numeric(is.na(GDP)), .after= GDP) |>
  group_by(Country) |>
  summarise(N_GDP_missing = sum(GDP_missing))
  
```

```{r}
# Average Development Level

d_gdp |>
  group_by(Country) |>
  summarise(GDP_average = mean(GDP, na.rm = TRUE))

```

```{r}
#GDP growth and GDP per capita growth: 2019 over 1984

d_gdp |>
  filter(Year >= 1984 & Year <= 2019) |>
  group_by(Country) |>
  arrange(Year) |>
  summarise(GDP_growth_2019_1984 = (last(GDP) - first(GDP)) / first(GDP),
            GDPpc_growth_2019_1984 = (last(GDP_per_capita) - first(GDP_per_capita)) / first(GDP_per_capita)) |>
  ungroup() |>
  arrange(Country)
```

# Homework

### Step 1 **Coodbook Lookup**

```{r}
library(tidyverse)
```

```{r}
d <- read_csv("/Users/kamalpathak/Desktop/Personal R /POLI3148_DaSPPA_Portfolio/_DataPublic_/vdem/1984_2022/vdem_1984_2022_external.csv")
```

Question: How unequal is the level of education achieved by the population aged 15 years and older?

Citation: Clio Infra (clio-infra.eu), Years: 1984-2010

Question: What is the average years of education among citizens older than 15? Clarification: The Average years of education in the total population aged 15 years and older.

Citation: Clio Infra (clio-infra.eu) Years: 1984-2022

### Step 2 **Subset by columns**

```{r}
names(d)
```

```{r}
d_edu <- d |> 
  select(country_name, , year, e_peaveduc, e_peedgini)

d_edu
```

```{r}
d_edu <- d_edu |>
  rename("Edu_15" = "e_peaveduc", "Edu_Ineq" = "e_peedgini",
         "Country" = "country_name", 
         "Year" = "year")

d_edu
```

### Step 3 **Subset by rows**

```{r}
d_edu |> slice_max(order_by = Edu_15, n = 5)


```

```{r}

d_edu |> slice_min(order_by = Edu_Ineq, n = 5)

```

### Step 4 **Summarize the data**

```{r}
# Data Availability

d_edu |> 

  #Create a column that indicates whether the value is missing
  mutate(Edu_missing = as.numeric(is.na(Edu_15)), .after= Edu_15) |>
  group_by(Country, Year) |>
  summarise(N_Edu_missing = sum(Edu_missing))
```

```{r}
#Average level of education quality from 1984 to 2022


d_edu |>
  summarise(
    edu_average = mean(Edu_15, na.rm=TRUE),
    edu_ineq_average = mean(Edu_Ineq, na.rm=TRUE)
)


```

```{r}
#Change of education quality from 1984 to 2022
d_edu |>
  group_by(Country) |>
  arrange(Year) |>
  mutate(Edu_over_1984 = Edu_15 / first(Edu_15)) |>
  ungroup() |>
  arrange(Country, Year)


```

```{r}
# The average education quality by country
average_edu <- d_edu |>
  group_by(Country) |>
  summarise(edu_average = mean(Edu_15, na.rm = TRUE)) |>
  arrange(desc(edu_average))

# The country with the best average education quality
best_country <- average_edu|>
  slice(1)

```

### Step 4B

```{r}
# Eduaction average quality 
average_edu <- d_edu |>
  group_by(Country) |>
  summarise(edu_average = mean(Edu_15, na.rm = TRUE)) |>
  arrange((edu_average))

# The country with the best average education quality
worst_country <- average_edu|>
  slice(-1)


# Its Niger ( WORST EDU quality ) 
```
