```{r}
2+ 2

log(1) # 0

log10(1) # 0

sqrt(16) # 4

8^2.  # 64

```

# Lesson 2 

### 2.2 Objects in R \-\-\--

```{r}
## Numeric object 

my_obj <- 48

my_obj <- 36

my_obj

print(my_obj) # 36
```

###  Character objects 

```{r}
obj_character <- "R is cool"

obj_character_2 <- 'R is cool'

obj_character_3 <- "R's cool"

print(obj_character_3)
# "R's cool"
```

## \# Remove objects 

```{r}
# Remove a single object

rm(my_obj)

# Remove all the objects 
rm(list = ls())
```

## \# 2.3 Vector 

#### Basic functions applied to vectors 

```{r}
## Basic functions applied to vectors 

my_vec <- c(2, 3, 1, 6, 4, 3, 3, 7)

my_vec

mean(my_vec)

var(my_vec) # variance

sd(my_vec) 

length(my_vec) # how many chrcters
```

#### Create sequences

```{r}
## Create sequences ----

#A sequence of numbers from 1 to 10.
my_seq <- 1:10


my_seq_revers <- 10:1


# A sequence of numbers starting from 1 and ending at 5,with increments of 0.5 between each number.
my_seq_advanced <- seq(from = 1, to = 5, by = 0.5)


my_seq_advanced <- seq(from = 0, to = 3, by = 0.3)

my_seq_advanced
```

#### Create repetition

```{r}

#A sequence of the number 2 repeated 10 times.
#2 2 2 2 2 2 2 2 2 2
my_seq_rep <- rep(2, times = 10)


my_seq_rep


# A sequence of numbers from 1 to 5 repeated 10 times.
my_seq_rep <- rep(1:5, times = 10)

my_seq_rep
```

#### Missing values

```{r}


temp <- c(1, 2, 3, NA)

temp

# NA is the indicator of missing values in R

mean(temp)

mean(temp, na.rm = TRUE)

# na.rm stands for NA remove

#------------------------------


# Getting Help ----

?mean
```

###  Chapter 3: Data in R 

#### Numeric 

```{r}


obj_numeric <- 4.3

class(obj_numeric) #numeric

typeof(obj_numeric) # double

is.numeric(obj_numeric) # TRUE
```

#### Integer 

```{r}

obj_numeric_2 <- 4L

class(obj_numeric_2) # integer

typeof(obj_numeric_2)

is.integer(obj_numeric_2)

```

####  Logical 

```{r}
obj_logical <- TRUE

obj_logical <- T

obj_logical <- FALSE

obj_logical <- F

is.logical(obj_logical)
```

####  Character 

```{r}
obj_character <- "R is cool"

is.character(obj_character)


```

#### Coersion 

```{r}
obj_to_change <- c("1", "3", "missing", "4", "5")

mean(obj_to_change)

obj_changed <- as.numeric(obj_to_change)

obj_changed # 1  3 NA  4  5
```

####  List 

```{r}
list_1 <- list(

c("black", "orange", "yellow"),

c(T, F),

matrix(1:6, nrow = 3)

)

list_1

list_1[[2]]

list_2 <- list(

color = c("black", "orange", "yellow"),

truth = c(T, F),

mat = matrix(1:6, nrow = 3)

)

list_2

list_2$color

list_3 <- list(

color = c("black", "orange", "yellow"),

truth = c(T, F, T)

)

as.data.frame(list_3)
```

#### Explaination of what's going on with list

#### \<!\--# list_1 is like a bag with three things inside: a list of colors (black, orange, yellow), a list of true and false values, and a grid of numbers arranged in rows and columns.

list_1[[2]] means taking out the second thing from the bag, which is the list of true and false values.

list_2 is another bag, but this time the things inside have names. It has a list of colors (black, orange, yellow), a list of true and false values, and a grid of numbers arranged in rows and columns.

list_2\$color means taking out the thing named "color" from the bag, which is the list of colors.

list_3 is a bag with two things inside: a list of colors (black, orange, yellow), and a list of true and false values.

as.data.frame(list_3) is like rearranging the things in the bag into a table-like format.

\--\>

####  Data Frame

```{r}
p.height <- c(180, 155, 160, 167, 181)

p.weight <- c(65, 60, 52, 58, 70)

p.names <- c("Joanna", "Charlotte", "Helen", "karen", "Amy")

dataf <- data.frame(height = p.height,

weight = p.weight,

names = p.names)

summary(dataf)

names(dataf)

dim(dataf)

summary(dataf$height)

summary(dataf$weight)

summary(dataf$names)

table(dataf$names)
```

#### Explainations about the above data frame 

##### Imagine we have a group of 5 people. We know their heights, weights, and names.

##### We can write down their heights in a list: 180, 155, 160, 167, 181.

##### We can also write down their weights: 65, 60, 52, 58, 70.

##### And we can write down their names: Joanna, Charlotte, Helen, Karen, Amy.

##### Now, we can put all this information together in a special table called a data frame.

##### The data frame is like a big table where each person has their own row, and each piece of information (height, weight, name) has its own column.

##### We can summarize this table to get some information about the whole group. 

##### For example, we can see the minimum and maximum height and weight, the average height and weight, and some other useful statistics.

##### We can also see the names of the columns in our table, which are "height", "weight", and "names".

##### We can know how many rows and columns our table has by checking its dimensions.

##### We can also get specific summaries for each column. For example, we can see the minimum and maximum height, weight, and the names.

##### 

Lastly, we can count how many times each name appears in our table using the "table" function.

### Get to specific locations of a data frame

```{r}
### Row 1

dataf[1, ]

### Column 2

dataf[, 2]

### Cell at row 1, column 2

dataf[1, 2]
```
